### 2018-03-17 paper/4-build.tex

\begin{table}[h]
\smaller
\centering
\begin{tabular}{l||c|c|c}
\hline
Property           & Topological \S\ref{sec-topological} & Reordering \S\ref{sec-reordering} & Recursive \S\ref{sec-recursive}    \\\hline
\hline
Dirty bit \S\ref{sec-dirty-bit}                    & \Make    & \Excel                   & Approximate \Shake  \\\hline
Verifying trace \S\ref{sec-verifying-traces}       & \Ninja   & Traced \Excel            & \Shake                      \\\hline
Constructive trace \S\ref{sec-constructive-traces} & \Bazel   & Cloud \Excel             & Cloud \Shake                \\\hline
\end{tabular}
\vspace{0.5mm}
\caption{Build systems \`a la Carte.\label{tab-build-systems}}
\end{table}

>>>>

\begin{table}[h]
\smaller
\centering
\begin{tabular}{l||c|c|c}
\hline
Property           & Topological \S\ref{sec-topological} & Reordering \S\ref{sec-reordering} & Recursive \S\ref{sec-recursive}    \\\hline
\hline
Dirty bit \S\ref{sec-dirty-bit}                    & \Make    & \Excel                   & Approximate \Shake*  \\\hline
Verifying trace \S\ref{sec-verifying-traces}       & \Ninja   & Traced \Excel*            & \Shake                      \\\hline
Constructive trace \S\ref{sec-constructive-traces} & \Bazel   & Cloud \Excel*             & Cloud \Shake*                \\\hline
\end{tabular}
\vspace{0.5mm}
\caption{Build systems \`a la Carte, systems marked * are hypothetical systems that do not currently exist.\label{tab-build-systems}}
\end{table}

### 2018-03-17 paper/5-implementations.tex
Delete some old commented out stuff

% Our code comprises a framework, utility functions implementing the
% various dependency orderings as per~\S\ref{sec-dependency-orderings}, and
% actual implementations.

% \subsection{Framework}

% todo{SLPJ}{I think this entire section, including the mysterious M, can be deleted.  Or is
% there anything that remains?}

% \begin{figure}
% \begin{minted}{haskell}
% type M i k v r = State (i, k -> v, ...) r
% runM :: Default i => M i k v a -> Maybe i -> (k -> v) -> (i, k -> v)

% getStore :: k -> M i k v v
% putStore :: k -> v -> M i k v ()
% getInfo :: M i k v i
% modifyInfo :: (i -> i) -> M i k v ()
% \end{minted}
% Note: we have omitted the context on \hs{k}, things like \hs{Eq} (to implement the store).
% \caption{API with which to implement build systems}
% \label{fig-M-api}
% \end{figure}

% We use \hs{runM} to switch between the pure world of \hs{Build} and a state monad \hs{M} which lets us more precisely express the ideas behind the build system without threading the state manually - the types of important operations are given in Figure~\ref{fig-M-api}. We provide functions like \hs{getStore} to load from the store and \hs{putStore} to write to the store. We'll introduce the other functions in in the monad \hs{M} as we need them while exploring the build systems.

@@

\begin{enumerate}
\item Any system using verifying traces, e.g. \Shake, is unlikely to see significant benefit from storing more than one \hs{Trace} per key\footnote{There is a small chance of a benefit if the dependencies change but the result does not, and then the dependencies change back to what they were before.}. Therefore, such systems can store \hs{Map k (Trace k v)}, where the initial \hs{k} is the \hs{key} field of \hs{Trace}.
\item Any system using \hs{Applicative} dependencies can omit the dependency keys from the \hs{Trace} as they can be recovered from the \hs{key} field.
\item Any \hs{Applicative} system storing constructive traces, e.g. \Bazel, can index directly from the key and results to the output result - i.e. \hs{Map (k, [Hash v]) (Hash v)}. Importantly, assuming the traces are stored on a central server, the client can compute the key and the hashes of its dependencies, then make a single call to the server to retrieve the result hash. In this formulation we have removed the possibility for a single key/dependency state to map to multiple different hashes, e.g. on a non-deterministic build - something \Bazel already prohibits which is discussed more in \S\ref{sec-non-determinism}.
\item Finally, a \hs{Monad} system with constructive traces can be efficiently encoded as \hs{Map k (Choice k v)}, assuming a definition of \hs{Choice} as:
\begin{minted}{haskell}
data Choice k v = Choice k (Map (Hash v) Choice)
                | Result (Hash v)
\end{minted}
Here the \hs{Choice} effectively encodes a tree, asking successive questions about keys, and taking different branches based on the resulting value, until it reaches a final result. Implementing this structure over client-server communication requires either a chatty interface with lots of round-trips per \hs{Choice} step, or sending over a part of the tree that is not subsequently explored.
\end{enumerate}

>>>>

@@

\begin{enumerate}
\item Any system using verifying traces, e.g. \Shake, is unlikely to see significant benefit from storing more than one \hs{Trace} per key\footnote{There is a small chance of a benefit if the dependencies change but the result does not, and then the dependencies change back to what they were before.}. Therefore, such systems can store \hs{Map k (Trace k v)}, where the initial \hs{k} is the \hs{key} field of \hs{Trace}.
\item Any system using \hs{Applicative} dependencies can omit the dependency keys from the \hs{Trace} as they can be recovered from the \hs{key} field.
\item Any \hs{Applicative} system storing constructive traces, e.g. \Bazel, can index directly from the key and results to the output result - i.e. \hs{Map (k, [Hash v]) (Hash v)}. Importantly, assuming the traces are stored on a central server, the client can compute the key and the hashes of its dependencies, then make a single call to the server to retrieve the result hash. In this formulation we have removed the possibility for a single key/dependency state to map to multiple different hashes, e.g. on a non-deterministic build - something \Bazel already prohibits which is discussed more in \S\ref{sec-non-determinism}.
\item Finally, a \hs{Monad} system with constructive traces can be represented as \hs{Map k (Choice k v)}, assuming a definition of \hs{Choice} as:
\begin{minted}{haskell}
data Choice k v = Choice k (Map (Hash v) Choice)
                | Result (Hash v)
\end{minted}
Here the \hs{Choice} effectively encodes a tree, asking successive questions about keys, and taking different branches based on the resulting value, until it reaches a final result. Implementing this structure over client-server communication requires either a chatty interface with lots of round-trips per \hs{Choice} step, or sending over a part of the tree that is not subsequently explored.
\end{enumerate}

### 2018-03-17 paper/6-engineering.tex
4-build

\todo{NM}{Isn't ``early aborting'' automatically handled by short-circuiting
withing monads, i.e. \hs{Nothing} or \hs{Left e} cancel out all subsequent
computation?}

@@

\item[Non-determinism] Some tasks are \emph{non-determinstic}, producing a
result from a possible set. As an example, GHC when compiled using parallelism
can change the order in which unique variables are obtained from the supply,
producing different but semantically identical results.

@@

\todo{NM}{I cited GVFS above, and could not find a proper reference to Google's
internal equivalent. There are some mentions online of Piper, but not citable.}

>>>>

% Q: {Isn't ``early aborting'' automatically handled by short-circuiting
% withing monads, i.e. \hs{Nothing} or \hs{Left e} cancel out all subsequent
% computation?}
% NM says: The build system could chose an early abort monad, but the tasks are free in the monad, so you can't get the encoding to cross between them.

@@

\item[Non-determinism] Some tasks are \emph{non-deterministic}, producing a
result from a possible set. As an example, GHC when compiled using parallelism
can change the order in which unique variables are obtained from the supply,
producing different but semantically identical results.

@@
